---
title: "Olympics notebook"
author: "Alexandre Facon-Bocquet"
---

# Load the data

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(plotly)
```

```{r}
setwd("/Users/alexandrefacon-bocquet/Documents/olympics")
olympic_data = read_csv("olympic_medals_tab.csv")
```

```{r}

olympic_data
```

# We are cleaning our data

removing the `athlete_url` column

```{r}
cleaned_olympics = subset(olympic_data, select = -c(athlete_url) )
cleaned_olympics
```

completing the NA values of the `athlete_full_name` column

-   checking all NA in the `athlete_full_name`

```{r}
cleaned_olympics[is.na(cleaned_olympics$athlete_full_name),]
```

-   we could see that all the NA names are related to a GameTeam participant_type. So we have to put a "Team" tag

```{r}
cleaned_olympics[is.na(cleaned_olympics$athlete_full_name),"athlete_full_name"] <- "Team Medal"
cleaned_olympics
```

We are checking nothing missing in the data set

```{r}
str(is.na(olympic_data))
```

Everything's good, we can go further in the analyse

# Visualization

## Medals raking

we could show how many medal each country have. First we count all the medals by countries et medal types.

```{r}
country_sum <- cleaned_olympics %>%
  group_by(country_name, medal_type) %>%
  summarise(
    n_medal = n()
  )
```

Now we can plot the bar graph of the medals repartition.

```{r}
library(plotly)

top100_countries <- country_sum %>%
  group_by(country_name) %>%
  summarize(total_medals = sum(n_medal)) %>%
  slice_max(order_by = total_medals, n = 100) # 154 pays

g <- ggplot(country_sum %>%
  filter(country_name %in% top100_countries$country_name),
  aes(x = reorder(country_name, n_medal), y = n_medal, fill = medal_type)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 100 Countries by Total Medals",
       x = "Country",
       y = "Total Medals") +
  theme_minimal()

g <- ggplotly(g) #, dynamicTicks = TRUE)
g <- g %>% layout(dragmode = "pan", hoverinfo = "n_medal")
g
```

## Olympics map

Let's create a map of all Olympics localisations. In a first time we create a data frame where we include localisations et editions' informations. Then we create a map that plot all the localisations assigned with there informations.

```{r}
olympic_data = read_csv("olympic_medals_tab.csv")

localisation <- olympic_data
localisation$season = paste(localisation$year, localisation$edition, sep = " ")

localisation <- localisation %>%
  select(city_host, season, lat, long) %>%
  filter(!duplicated(season))
  

library(leaflet)
library(leaflet.extras)

myIcon <- makeIcon(
  iconUrl = "Olympics-Logo.png",
  iconWidth = 50, #iconHeight = 95,
  iconAnchorX = 15, iconAnchorY = 24,
)

# Create a base map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%  
  #addTiles() %>%
  addMarkers(lng = localisation$long, 
             lat = localisation$lat, 
             popup = paste0("<strong>", localisation$city_host, "</strong><br>","Edition: ", localisation$season),
             clusterOptions = markerClusterOptions(),
             icon = myIcon)
```
